---
title: "Introduction"
---

```{r}
#| label: setup
#| include: false
source("../R/_setup.R")
source("../R/parameters.R")
```

## Clinical Context

### Long COVID Overview

Long COVID, also known as Post-Acute Sequelae of SARS-CoV-2 (PASC), affects a significant proportion of individuals following COVID-19 infection. Common symptoms include:

-   Persistent fatigue
-   Cognitive dysfunction ("brain fog")
-   Sleep disturbances
-   Reduced exercise tolerance

These symptoms can persist for months or years, significantly impacting quality of life and functional capacity.

### Taurine as a Potential Intervention

Taurine is a conditionally essential amino acid with multiple physiological roles:

-   **Antioxidant properties**: May reduce oxidative stress associated with Long COVID
-   **Neuroprotective effects**: Potential to improve cognitive function
-   **Anti-inflammatory action**: May help address systemic inflammation
-   **Energy metabolism**: Could improve mitochondrial function and reduce fatigue

This trial investigates whether taurine supplementation can improve cognitive function and reduce fatigue in Long COVID patients.

## Co-Primary Outcomes

This trial uses two co-primary endpoints, requiring both to show benefit for trial success:

### TMT B/A Ratio (Trail Making Test)

The Trail Making Test is a neuropsychological measure of cognitive function:

-   **What it measures**: Executive function, processing speed, mental flexibility
-   **Scoring**: Ratio of Part B to Part A completion times
-   **Interpretation**: Lower values indicate better cognitive function
-   **Clinical range**: 0.9 - 5.0
-   **MCID**: 0.5 units

```{r}
#| label: tmt-params
cat(sprintf("Baseline parameters:\n"))
cat(sprintf("  Mean: %.2f\n", outcomes$tmt$mean))
cat(sprintf("  SD: %.2f\n", outcomes$tmt$sd))
cat(sprintf("  Adjusted mean (for LC population): %.2f\n", get_adjusted_mean("tmt")))
```

### MFIS (Modified Fatigue Impact Scale)

The MFIS is a validated measure of fatigue severity:

-   **What it measures**: Physical, cognitive, and psychosocial fatigue
-   **Scoring**: Sum of 21 items, each rated 0-4
-   **Interpretation**: Lower values indicate less fatigue
-   **Clinical range**: 0 - 84
-   **MCID**: 10 points

```{r}
#| label: mfis-params
cat(sprintf("Baseline parameters:\n"))
cat(sprintf("  Mean: %.1f\n", outcomes$mfis$mean))
cat(sprintf("  SD: %.1f\n", outcomes$mfis$sd))
cat(sprintf("  Adjusted mean (for LC population): %.1f\n", get_adjusted_mean("mfis")))
```

## Trial Design

### Randomization

Participants are randomized in a **2:1 ratio** (Treatment:Control):

-   Treatment group: 67% of participants
-   Control group: 33% of participants

This unequal allocation maximizes information about the treatment effect while maintaining adequate control group size.

### Sample Size Considerations

The primary goal of this simulation study is to determine the sample size required to achieve:

-   **90% power** to detect clinically meaningful effects on both outcomes
-   Using a **Bayesian decision criterion** of 95% posterior probability

### Follow-up Duration

Participants are assessed at:

-   **Baseline**: Pre-randomization assessment
-   **3 months**: Primary endpoint assessment

## Success Criteria

### Primary Decision Rule

The trial will be considered successful if:

$$P(\text{TMT effect} < 0 \,|\, \text{data}) \geq 0.95 \quad \text{AND} \quad P(\text{MFIS effect} < 0 \,|\, \text{data}) \geq 0.95$$

This requires strong evidence of benefit for **both** co-primary outcomes.

### Interim Stopping Rules

The trial includes planned interim analyses for:

-   **Early efficacy**: Stop if overwhelming evidence of benefit for both outcomes
-   **Futility**: Stop if either outcome shows negligible probability of benefit

These stopping rules are detailed in @sec-interim-analysis.

## Simulation Objectives

This simulation study aims to:

1.  Estimate the required sample size for 90% power
2.  Evaluate the operating characteristics of the interim stopping rules
3.  Calculate the expected sample size under various scenarios
4.  Assess sensitivity to prior specifications and effect size assumptions