---
title: "Operating Characteristics"
---

```{r}
#| label: setup
#| include: false
source("../R/_setup.R")
source("../R/parameters.R")
source("../R/priors.R")
source("../R/simulate_data.R")
source("../R/fit_model.R")
source("../R/power_analysis.R")
source("../R/type1_error.R")
```

## Overview

This chapter presents the complete operating characteristics of the Bayesian design for the taurine supplementation trial:

1. **Conditional Power**: Probability of success at fixed (design) effect sizes
2. **Assurance**: Expected probability of success averaging over design prior
3. **Type I Error**: False positive rate under null hypothesis
4. **Required Sample Size**: N needed to achieve target metrics

These metrics provide a comprehensive picture of the design's performance for regulatory review.

## Conditional Power Analysis

### Power at Design Effect Size

Conditional power is calculated at the design effect sizes specified in `parameters.R`:

```{r}
#| label: design-effects
#| echo: false
cat("Design Effect Sizes (gamma_treat):\n")
cat(sprintf("  TMT: %.3f (standardized)\n", true_effects$tmt$gamma_treat))
cat(sprintf("  MFIS: %.3f (standardized)\n", true_effects$mfis$gamma_treat))
```

### Power Curve Computation

```{r}
#| label: power-curve
#| eval: false
#| echo: true

# Compile Stan model
model <- compile_model()

# Estimate power curve across sample sizes
power_results <- estimate_power_curve(
  sample_sizes = sim_params$sample_sizes,
  model = model,
  n_reps = sim_params$n_reps,
  parallel = FALSE
)

# Save results
saveRDS(power_results, "data/cached_results/power_results.rds")
```

```{r}
#| label: power-display
#| eval: false

# Load cached results
power_results <- readRDS("data/cached_results/power_results.rds")

# Display power summary
print_power_summary(power_results)
```

### Power Curve Visualization

```{r}
#| label: power-plot
#| eval: false
#| fig.cap: "Conditional power curve with 95% confidence intervals"

ggplot(power_results, aes(x = n, y = power)) +
  geom_line(linewidth = 1, color = "#0072B2") +
  geom_ribbon(aes(ymin = lower_ci, ymax = upper_ci),
              alpha = 0.2, fill = "#0072B2") +
  geom_point(size = 2, color = "#0072B2") +
  geom_hline(yintercept = 0.8, linetype = "dashed", alpha = 0.5) +
  geom_hline(yintercept = 0.9, linetype = "dashed", alpha = 0.5) +
  annotate("text", x = max(sim_params$sample_sizes), y = 0.82,
           label = "80% power", hjust = 1, size = 3) +
  annotate("text", x = max(sim_params$sample_sizes), y = 0.92,
           label = "90% power", hjust = 1, size = 3) +
  scale_y_continuous(limits = c(0, 1), labels = scales::percent) +
  labs(
    x = "Total Sample Size (N)",
    y = "Power",
    title = "Conditional Power at Design Effect Size",
    subtitle = sprintf("TMT effect = %.2f, MFIS effect = %.2f (standardized)",
                      true_effects$tmt$gamma_treat, true_effects$mfis$gamma_treat)
  ) +
  theme_minimal(base_size = 12)
```

::: {.callout-note}
## Placeholder
Run the power analysis code to generate results. Results are cached in `data/cached_results/` for subsequent renders.
:::

## Assurance Analysis

Assurance accounts for uncertainty in the true treatment effect by averaging power over the design prior. This provides a more honest assessment of trial success probability.

### Assurance Computation

```{r}
#| label: assurance
#| eval: false
#| echo: true

# Create design prior
design_prior <- create_combined_design_prior(
  tmt_effect = -0.10,
  tmt_sd = 0.05,
  mfis_effect = -0.20,
  mfis_sd = 0.10
)

# Estimate assurance curve
assurance_results <- estimate_assurance_curve(
  sample_sizes = sim_params$sample_sizes,
  model = model,
  design_prior = design_prior,
  n_reps = sim_params$n_reps
)

# Save results
saveRDS(assurance_results, "data/cached_results/assurance_results.rds")
```

```{r}
#| label: assurance-display
#| eval: false

# Load cached results
assurance_results <- readRDS("data/cached_results/assurance_results.rds")

# Display assurance summary
print_assurance_summary(assurance_results)
```

::: {.callout-important}
## Power vs. Assurance
Assurance is typically lower than conditional power because it accounts for the possibility that the true effect may be smaller than the design assumption. This makes assurance a more conservative and honest metric for planning.
:::

## Type I Error

Type I error (false positive rate) is estimated by simulating trials under the null hypothesis (no treatment effect).

### Type I Error Estimation

```{r}
#| label: type1-error
#| eval: false
#| echo: true

# Estimate Type I error at key sample sizes
type1_results <- estimate_type1_curve(
  sample_sizes = sim_params$sample_sizes,
  model = model,
  n_reps = 500,  # More replications for stable estimate
  decision_threshold = sim_params$efficacy_threshold
)

# Save results
saveRDS(type1_results, "data/cached_results/type1_results.rds")
```

### Type I Error Summary

```{r}
#| label: type1-display
#| eval: false

type1_results <- readRDS("data/cached_results/type1_results.rds")

cat("Type I Error by Sample Size:\n")
cat(sprintf("Decision threshold: %.2f\n\n", sim_params$efficacy_threshold))
for (i in seq_len(nrow(type1_results))) {
  cat(sprintf("  N = %3d: Type I Error = %.4f [%.4f, %.4f]\n",
              type1_results$n[i],
              type1_results$type1_error[i],
              type1_results$lower_ci[i],
              type1_results$upper_ci[i]))
}
```

::: {.callout-tip}
## Regulatory Context
For a 0.95 posterior probability threshold with weakly informative priors, the Type I error is typically well below 0.05, providing strong error control.
:::

## Required Sample Size

### For Target Power

```{r}
#| label: required-n-power
#| eval: false
#| echo: true

# Estimate required N for 90% power
required_n_power <- estimate_required_n(power_results, target_power = 0.90)

cat("Required Sample Size for 90% Power:\n")
cat(sprintf("  Point estimate: %.0f\n", required_n_power$required_n))
cat(sprintf("  95%% CI: [%.0f, %.0f]\n",
            required_n_power$lower_ci, required_n_power$upper_ci))
```

### For Target Assurance

```{r}
#| label: required-n-assurance
#| eval: false
#| echo: true

# Estimate required N for 80% assurance
required_n_assurance <- estimate_required_n_assurance(assurance_results, target_assurance = 0.80)

cat("Required Sample Size for 80% Assurance:\n")
cat(sprintf("  Point estimate: %.0f\n", required_n_assurance$required_n))
cat(sprintf("  95%% CI: [%.0f, %.0f]\n",
            required_n_assurance$lower_ci, required_n_assurance$upper_ci))
```

## Operating Characteristics Table

```{r}
#| label: oc-table
#| eval: false
#| tbl-cap: "Operating Characteristics by Sample Size"

# Create combined OC table
oc_table <- create_oc_table(power_results, assurance_results, type1_results)

knitr::kable(
  oc_table,
  digits = 3,
  col.names = c("N", "Power", "Lower", "Upper",
                "Assurance", "Lower", "Upper",
                "Type I Error", "Lower", "Upper"),
  align = "rrrrrrrrrr"
)
```

## Summary

| Metric | Target | Achieved | Sample Size |
|--------|--------|----------|-------------|
| Power | 90% | [Run analysis] | [Run analysis] |
| Assurance | 80% | [Run analysis] | [Run analysis] |
| Type I Error | ≤5% | [Run analysis] | N/A |

::: {.callout-note}
## Running the Analysis
To generate these results, run the power analysis scripts:
```r
source("R/_setup.R")
source("R/parameters.R")
source("R/priors.R")
source("R/simulate_data.R")
source("R/fit_model.R")
source("R/power_analysis.R")
source("R/type1_error.R")

model <- compile_model()
# Then run the computation blocks above
```
:::
