---
title: "Executive Summary"
---

```{r}
#| label: setup
#| include: false
source("R/_setup.R")
source("R/parameters.R")
```

## Study Overview

This report presents a Bayesian simulation-based sample size determination for a clinical trial investigating **taurine supplementation in Long COVID patients**. The trial uses two co-primary endpoints:

1. **TMT B/A Ratio** (Trail Making Test) - Cognitive function measure
2. **MFIS** (Modified Fatigue Impact Scale) - Fatigue severity measure

## Key Design Features

- **Randomization**: 2:1 (Treatment:Control)
- **Bayesian decision threshold**: 95% posterior probability of benefit
- **Interim analyses**: Every 60 participants after initial 120
- **Stopping rules**:
  - Efficacy: Both outcomes show P(benefit) > 0.95
  - Futility: Either outcome shows P(benefit) < 0.10

## Trial Parameters

```{r}
#| label: parameters-summary
print_parameters()
```

## Report Structure

This report contains the following chapters:

1. **Introduction**: Clinical context and study rationale
2. **Methods**: Bayesian model specification and decision rules
3. **Simulation**: Data generation methodology and validation
4. **Power Analysis**: Sample size determination for 90% power
5. **Interim Analysis**: Adaptive design with stopping rules
6. **Results**: Final recommendations
7. **Sensitivity**: Prior and effect size sensitivity analyses

## Quick Results

::: {.callout-note}
## Note
The detailed simulation results are presented in subsequent chapters. This executive summary will be updated with key findings after running the full analysis pipeline.
:::

To run the full analysis, execute:

```r
source("R/_setup.R")
source("R/parameters.R")
source("R/simulate_data.R")
source("R/fit_model.R")
source("R/power_analysis.R")
source("R/interim_analysis.R")
source("R/visualization.R")

# Compile model
model <- compile_model()

# Run power analysis (computationally intensive)
power_results <- estimate_power_curve(
  sample_sizes = sim_params$sample_sizes,
  n_reps = sim_params$n_reps,
  model = model,
  parallel = TRUE
)

# Estimate required N
required_n <- estimate_required_n(power_results)
print_power_summary(power_results, required_n)
```
